---
title: "Visual1: ggplot1"
author: "Ying Chen"
date: "9/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# only use this package for onething, but not too much
library(ggridges)
library(hexbin)

```

## base R has aweful plot, don't use it, on the right hand is the ggplot 
## has legend and axis are well labeled
## ggplot will be the function to use all the time for now
## 2013-2014 came out ggplot, 
## ggplot2/latice is another graphic function
## tidy data is very important!!
## grammer of graphics (gg)
## starting with dataset, features of graphic (axis, color) that you care and map to variables of data, geom to show 
## panels ... be consistent ... 


```{r load_data, cache=TRUE}
# national geographic and weather service data
# data are publically available
# download data from three stations
# daily weather information of the three stations
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    # .1 celc has to devided by 10
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## create a ggplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))

ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()

# alternative way of making this plot
# default is printing the plot
# Most time not using the style
#weather_df %>% filter(name =="CentralPark_NY")
#scaterplot = weather_df %>% 
#  ggplot(aes(x = timin, y = tmax)) + geom_point()
#scaterplot
#weather_df %>%
#  ggplot(aes(x = tmin, y = tmax)) + 
#  geom_point()

plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()

ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))

ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  # se: grey area not certain area, not very useful to set as true
  geom_smooth(se = FALSE)

# facet
# do not send multiple plots.
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)

# this is fine, but not very interesting
# color = name is global
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  #size = prcp, put bubbles accounts for prcp amont
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)

# alpha is the transparent level. .1 is not at all
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(tmax_fahr = tmax * (9 / 5) + 32,
         tmin_fahr = tmin * (9 / 5) + 32) %>% 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)

```

## why do 'aes' positions mater? se = false, no CIs
## define color at x, y level or geom level makes the plots looks different
## at geom level, the smooth color has only one color
## outside geom, it applies smooth line colors according to the names (three different smooth lines)
## can modify the graphics features quickly with little modification.

## some extra stuff

```{r extra}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE)

ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()

ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))

ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()

ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)


ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")

ggplot(weather_df, aes(x = name, y = tmax)) + geom_boxplot()

ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 4)

ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)

```


``` {r}
ggplot(weather_df, aes(x = prcp)) + 
  geom_density(aes(fill = name), alpha = .5) 

ggplot(weather_df, aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)


ggplot(weather_df, aes(y = prcp, x = name)) + 
  geom_boxplot() 

weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)

weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
