---
title: "Viz_EDA"
author: "Ying Chen"
date: "10/3/2019"
output: github_document
---
 
```{r EDA, include=FALSE}

library(tidyverse)
library(viridis)
library(ggridges)
library(hexbin)

knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE,
  fig = fig.width = 8,
  fig.height = 6,
  out.width = "90%")
```
 
### compute numerica data for EDA
### imported and coded correctly
### producing numerica summary
### groups that you are interested in. less specified. defining different groups. average responses acrossinng groups
### nested within other variables. 
### multiple layers: tx group, gender groups, plots, 
### measurement of center, mean, median, etc.
### variability: std, median abosolute deviation, n x group, boxplot, 

## tools: group_by(), summarize()
### can do a lot of different things
### does not change your dataset with grouping, added a layer ontop of the dataset
### group once created, it will be there and should be ungroup later. Keep in mind
# difficult to debug!!!

## EDA: lead to formal analysis not valuable later. not overtaken the primary analysis
## preplanned analysis, not random thing. stay with the primary analsis first if NS, go to 2ndary analysis. 


```{r}
weather_df =
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"),
                      date_min = "2016-01-01",
                      date_max = "2016-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY",
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, date, month, everything())
 
 
weather_df %>%
  group_by(name, month)
 
weather_df %>%
  group_by(month) %>%
  summarize(n = n())
 
weather_df %>%
  count(month, name = "n_days")
 
weather_df %>%
  pull(month) %>%
  table
 
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
 
weather_df %>%
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >=5 ~ "not_cold",
      TRUE     ~ ""
  )) %>%
  filter(name != "Waikiki_HA") %>%
  group_by(name, cold) %>%
  summarize(count = n())
 
weather_df %>%
  mutate(cold = case_when(
    tmax < 5 ~ "cold",
    tmax >=5 ~ "not_cold",
    TRUE     ~ ""
  )) %>%
  filter(name != "Waikiki_HA") %>%
  janitor::tabyl(name, cold)
 
```
 
##Standard statistical summaries are regularly computed in summarize() using functions like mean(), median(), var(), sd(), mad(), IQR(), min(), and max(). To use these, you indicate the variable to which they apply and include any additional arguments as necessary.
 
 
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))
 
weather_df %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax),
    median_tmax = median(tmax))
 
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
    geom_point() + geom_line() +
    theme(legend.position = "bottom")
 
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>%
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>%
  knitr::kable(digits = 1)
 
weather_df %>%
  group_by(name) %>%
  mutate(centered_tmax = tmax - mean(tmax)) %>%
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
    geom_point()
 
weather_df %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax))
 
weather_df %>%
  group_by(name, month) %>%
  filter(min_rank(tmax) < 2)
 
weather_df %>%
  group_by(name, month) %>%
  filter(min_rank(desc(tmax)) < 4)
 
```
 
 
### lags, are used to compare an observation to it’s previous value.
 
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax))
 
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax)) %>%
  summarize(temp_change_sd = sd(temp_change, na.rm = TRUE),
            temp_change_max = max(temp_change, na.rm = TRUE))
 
```
### Practices
 
```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)
 
pulse_data %>%
  group_by(visit) %>%
  summarize(
    mean_bdi = mean(bdi, na.rm = TRUE),
    median_bdi = median(bdi, na.rm = TRUE)) %>%
  knitr::kable(digits = 3)
 
pup_data =
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))
 
litter_data =
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
         day_of_tx = as.numeric(day_of_tx))
 
fas_data = left_join(pup_data, litter_data, by = "litter_number")
 
fas_data %>%
  group_by(dose, day_of_tx) %>%
  summarize(mean_pivot = mean(pd_pivot, na.rm = TRUE)) %>%
  filter(!is.na(dose)) %>%
  spread(key = dose, value = mean_pivot) %>%
  knitr::kable(digits = 3)
```
 
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
